Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-11-10 PAPER.tex
-Standardized eqref across whole paper -Standardized superscript alpha's and beta's on Q's instead of subscripts

\begin{equation}
\label{eq:qhat}
\hat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}_{\alpha}\textbf{R}\textbf{x}
\end{equation}

@@

In addition to foreground bias, noise bias can also be responsible for positive power spectrum detections. One example is noise bias arising if thermal noise is multiplied by itself. Every $21$ cm visibility measurement contains thermal noise that is comprised of receiver and sky noise. We expect this noise to be independent between antennas and thus we can beat it down (increase sensitivity) by integrating longer, using more baselines, etc. However, the squaring of noise occurs when cross-multiplying visibilities, which is shown by the two copies of $\textbf{x}$ in Equation \ref{eq:qhat}. If both copies of $\textbf{x}$ come from the same baseline and time, it can result in power spectrum measurements that are higher than those predicted by the thermal noise of the instrument. One way to avoid this type of noise bias is to avoid cross-multiplying data from the same baselines or days. This ensures that the two quantities that go into a measurement have separate noises that don't correlate with each other.

@@

\begin{equation}
\label{eq:Pin}
P_{in,\alpha} \propto \textbf{e}^{\dagger}\textbf{I}\textbf{Q}_{\alpha}\textbf{I}\textbf{e}
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{out,\alpha} \equiv \hat{\textbf{p}}_{e,\alpha} &=& \hat{\textbf{p}}_{r,\alpha}-\hat{\textbf{p}}_{x,\alpha} \nonumber \\
&\propto& \textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}\textbf{r} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}_{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

@@

One may wonder why $P_{out}$ cannot be computed simply as the weighted power spectrum of $\textbf{e}$ alone, namely $P_{out,\alpha} \propto \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}\textbf{e}$. Expanding Equation \ref{eq:sigloss}, we see that $P_{out}$ is comprised of multiple terms:

\begin{eqnarray}
\label{eq:crossterm}
P_{out,\alpha} &\propto& (\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}_{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&\propto& \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}\textbf{x} + \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}\textbf{e} + \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}\textbf{x} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}_{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

@@

\begin{eqnarray}
\label{eq:pout_expand}
P_{out, \alpha,\textbf{e} \gg \textbf{x}} &\propto& \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}\textbf{e} + \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}_{\alpha}\textbf{R}_{r}\textbf{x}.
\end{eqnarray}

We see that our naive expression for $P_{out}$ is the first term in Equation \ref{eq:pout_expand}, but there are also two additional terms. An initial assumption would be that the cross-terms that involve both $\textbf{e}$ and $\textbf{x}$ should be zero, since foregrounds and the cosmological signal are physically unrelated. In \citet{ali_et_al2015}, this assumption is used, and the first term in Equation \ref{eq:pout_expand} is directly compared to the input (unweighted) power spectrum to compute signal loss.

However, deeper investigations of these terms reveal that they contain non-negligible power. In fact, foreground modes and signal modes are anti-correlated on average, as derived and explained in \citet{switzer_et_al2015}. The spurious correlation of the two means that we cannot compute signal loss using a signal-only simulation, which would yield greater values for $P_{out}$ and thereby underestimate signal loss. Therefore, in our revised signal loss computation we use the full quantity for $P_{out}$ as defined in Equation \ref{eq:sigloss}, which subtracts the weighted power spectrum of the data from the weighted power spectrum of data plus EoR.

@@

\begin{eqnarray}
P_{in,\alpha} &\propto& \textbf{e}^{\dagger}\textbf{I}\textbf{Q}_{\alpha}\textbf{I}\textbf{e} \\
P_{out,\alpha} &\propto& \textbf{r}^{\dagger}\textbf{C}_{r}^{-1}\textbf{Q}_{\alpha}\textbf{C}_{r}^{-1}\textbf{r} - \textbf{x}^{\dagger}\textbf{C}_{x}^{-1}\textbf{Q}_{\alpha}\textbf{C}_{x}^{-1}\textbf{x} 
\end{eqnarray}

Our signal loss transfer function for one $k$ value is shown in Figure \ref{fig:sigloss_transfercurve} and compares the weighted case ($\textbf{R} = \textbf{C}^{-1}$, left plot) with an unweighted case ($\textbf{R} = \textbf{I}$, right plot). The figures depict a smoothed-out distribution (via kernel density estimators) of our input and output power spectra. To sample the distribution, we run $40$ total injection levels, where the amplitude of \textbf{e} is increased in even increments in log-space and ranges from $P_{in} \sim -10^{13} \, mK^{2} (h^{-1} Mpc)^{3}$ (where one copy of \textbf{e} in Equation \ref{eq:Pin} is negative and the other is positive) to $P_{in} \sim10^{13} \, mK^{2} (h^{-1} Mpc)^{3}$. For each injection, we perform $20$ bootstraps (over baselines, as explained in Section \ref{sec:Boot}), resulting in a total of $800$ data points on our $P_{in}$ vs. $P_{out}$ grid. It is noted that the transfer function is symmetric for negative and positive $P_{in}$'s, and so we fold all the values into positive ones to increase our signal-to-noise.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve.png}
	\caption{Signal loss transfer functions showing the relationship of $P_{in}$ and $P_{out}$, as defined by Equations \ref{eq:Pin} and \ref{eq:sigloss}. Smoothed out distributions (via kernel density estimators) are shown as colored heat-maps for the cases of inverse covariance weighted PAPER-64 data (left) and unweighted data (right). The solid black diagonal line marks a perfect unity mapping, and the solid grey horizontal line denotes the peak of $\hat{\textbf{p}}$, the data distribution. From these plots, it is clear that inverse covariance weighting results in $\sim3$ orders of magnitude of signal loss whereas the unweighted case does not exhibit loss.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

One peculiar aspect of Figure \ref{fig:sigloss_transfercurve} is the fact that at low $P_{in}$ values it appears that we can have signal gain ($P_{out} > P_{in}$). This is unphysical in nature but caused due to the cross-terms dominating in Equation \ref{eq:pout_expand} once $\textbf{e}$ becomes small. As $\textbf{e}$ increases, we move into a regime where $P_{out} \sim P_{in}$, and then eventually into a regime where $P_{out} < P_{in}$ when $\textbf{e}$ is large enough to destroyed if weighting the data using itself. Although we only show figures for one $k$ value, we note that the shape of the transfer curve is nearly identical for all $k$'s (though we treat each $k$ separately).

@@

The first component is merely the fact that the values of the cross-terms in Equation \ref{eq:pout_expand} change for different EoR signals, $\textbf{e}$. Since we draw different random $\textbf{e}$'s for every bootstrap, there is a (small) range of $P_{out}$ values that could result for the same $P_{in}$. This randomness is what causes the spread seen in the unweighted case (right plot of Figure \ref{fig:sigloss_transfercurve}), as well as most of the spread in the weighted case (left plot).

@@

Equation \ref{eq:sense} represents the most sensitive estimate for $\hat{\textbf{p}}(k)$ given our observing parameters. An additional factor of $\sqrt{2}$ is gained in sensitivity when folding our power spectra into $\Delta^{2}(k)$, due to averaging together positive and negative $k$'s.

@@

We fringe-rate filter the noise in the same way as PAPER-64 data, using an optimal fringe-rate filter. Power spectrum results for the noise simulation are shown in Figure \ref{fig:ps_noise}, where the black and grey points represent positive and negative power spectrum values, respectively (with $2\sigma$ error bars and weighting matrix $\hat{\textbf{C}}_{eff}$), the dashed blue line represents the unweighted power spectrum, and the solid green line denotes our $2\sigma$ theoretical noise prediction as calculated by Equation \ref{eq:sense}. All three are in agreement, validating our analytical thermal noise calculation.

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_noise_reg.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_noise_reg.png}
	\caption{Power spectrum estimate for a noise simulation that mimics the noise level of PAPER-64 data. The weighted power spectrum points and $2\sigma$ errors are show in black and grey (positive and negative values), where we regularize $\hat{\textbf{C}}$ by setting $\hat{\textbf{C}} = \hat{\textbf{C}} * \textbf{I}$ to minimize signal loss. The dashed blue line is the unweighted power spectrum (also $2\sigma$ upper limit). The solid green line is the theoretical noise level prediction as calculated by Equation \ref{eq:sense}. All three estimates agree.}
	\label{fig:ps_noise}
\end{figure*}

@@

Tailoring window functions is one way to suppress foreground biases. As alluded to in Section \ref{sec:SiglossOverview}, we have a choice for the normalization matrix $\textbf{M}$ in Equation \ref{eq:phat}. For the analysis of PAPER-64 data, we compute $\textbf{M}$ using the Fisher matrix $\textbf{F}$, defined as:

\begin{equation}
\textbf{F}_{\alpha\beta} = \frac{1}{2} \text{tr} [\textbf{R}\textbf{Q}_{\alpha}\textbf{R}\textbf{Q}_{\beta} ]
\end{equation}

@@

In addition to mitigating foreground bias at high $k$'s, two other sources of bias that we actively suppress in the PAPER-64 analysis are noise bias associated with the squaring of thermal noise and noise bias from crosstalk. In order to avoid the former, we filter out certain cross-multiplications when forming $\hat{q}$ in Equation \ref{eq:qhat}. Namely, the PAPER-64 dataset is divided into two halves: even julian dates and odd julian dates. Our data vectors are then $\textbf{x}_{even, 1}$ for the `even' dataset and baseline group $1$, $\textbf{x}_{odd, 1}$ for the `odd' dataset and baseline group $1$, etc. We only form $\hat{q}$ when the two copies of $\textbf{x}$ come from different groups and baselines, never multiplying `even' with `'even', for example, in order to prevent the squaring of the same thermal noise.

@@

@@

The minimum-variance quadratic estimator $\hat{p}_\alpha$ for the $\alpha$th bandpower of the power spectrum is given by \acl{We need to decide whether the Q matrices have their Greek indices as upper or lower indices.} \acl{Also need to state in a footnote that for this section only we'll assume without loss of generality that the data are real}
\begin{equation}
\hat{p}_\alpha = \frac{1} {2 \F_{\alpha \alpha} }\x^t \C^{-1} \Q^{\alpha} \C^{-1} \x,
\end{equation}
where
\begin{equation}
F_{\alpha \alpha} \equiv \frac{1}{2} \textrm{tr} \left( \C^{-1} \Q^\alpha \C^{-1} \Q^\alpha \right)
\end{equation}
is the $\alpha$th diagonal element of the Fisher matrix. In our case, however, we do not have \emph{a priori} knowledge of the covariance matrix. Thus, we replace $\C$ with $\Chat$, its data-derived approximation. Our estimator then becomes
\begin{equation}
\label{eq:phatloss}
\hat{p}_\alpha^\textrm{loss} = \frac{1} {2 \F_{\alpha \alpha} }\x^t \Chat^{-1} \Q^{\alpha} \Chat^{-1} \x,
\end{equation}
where
\begin{equation}
\hat{F}_{\alpha \alpha} \equiv \frac{1}{2} \textrm{tr} \left( \Chat^{-1} \Q^\alpha \Chat^{-1} \Q^\alpha \right),
\end{equation}
with the label ``loss" to foreshadow the fact that this will be an estimator with signal loss (i.e., a multiplicative bias of less than unity). We will now provide an explicit demonstration of this by modeling the estimated covariance as
\begin{equation}
\label{eq:ChatDef}
\Chat = (1-\eta) \C + \eta \x \x^t,
\end{equation}
where $\eta$ is parameter quantifying our success at estimating the true covariance matrix. If $\eta = 0$, our covariance estimate has perfectly modeled the true covariance and $\Chat = \C$. On the other hand, if $\eta =1$, then our covariance estimate is based purely on the one realization of the covariance that is our actual data, and we would expect a high level of overfitting and signal loss.

Our strategy for computing the signal loss will be to insert Eq. \eqref{eq:ChatDef} \acl{Make sure we're consistent w.r.t. ``Equation" vs. ``Eq."} into Eq. \eqref{eq:phatloss} and to express the resulting estimator $\hat{p}_\alpha^\textrm{loss}$ in terms of $\hat{p}_\alpha$. We begin by expressing $\Chat^{-1}$ in terms of $\C^{-1}$ using the Woodbury identity so that

@@

where $\beta \equiv \eta g/[1+ \eta (g-1)]$ and $\varepsilon \equiv \beta / g$. To proceed, we take the ensemble average of $\hat{p}_\alpha^\textrm{loss}$ to understand its average statistical properties. Inspecting Eq. \eqref{eq:phatlossexpanded}, we see that its expectation value is a linear sum of terms of the form $\langle \beta^a h_\alpha^b h_{\alpha \alpha}^c \rangle$, where $a$, $b$, and $c$ are integers. Expanding $\beta$ in powers of $g$, we have

@@

and re-inserting this into Eq. \eqref{eq:betahh} yields

@@

With this result, taking the ensemble average of Eq. \eqref{eq:phatlossexpanded} is relatively straightforward. One simply replaces every copy of $\beta$ with $\beta_N \equiv \eta N / [1 + \eta (N-1)]$, every copy of $\varepsilon$ with $\varepsilon_N \equiv  \eta  / [1 + \eta (N-1)]$, every copy of $h_\alpha$ with $ \textrm{tr} (\C^{-1} \Q^\alpha )$, and every copy of $h_{\alpha \alpha}$ with $2F_{\alpha \alpha}$. Performing these substitutions and recalling \acl{make sure the reader will indeed be recalling at this point} that $\langle \hat{p}_\alpha \rangle = \langle h_\alpha \rangle / 2 F_{\alpha \alpha}$, we obtain

@@

In general, one's estimate of the covariance will not be expressible in the form given by Eq. \eqref{eq:ChatDef}. However, the qualitative point here is that the signal loss can be expressed as a multiplicative factor. Even if $\Chat$ is more complicated than the form used for this toy model, the same mathematical technique of expressing the estimator as a series of higher-point functions, reducing these functions (assuming Gaussianity) to two-point functions, and re-summing any power series will show that the signal loss is a multiplicative factor. \acl{I can express this last bit more eloquently, but it's late and I'm tired and need to eat dinner.}

>>>>

\begin{equation}
\label{eq:qhat}
\hat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}
\end{equation}

@@

In addition to foreground bias, noise bias can also be responsible for positive power spectrum detections. One example is noise bias arising if thermal noise is multiplied by itself. Every $21$ cm visibility measurement contains thermal noise that is comprised of receiver and sky noise. We expect this noise to be independent between antennas and thus we can beat it down (increase sensitivity) by integrating longer, using more baselines, etc. However, the squaring of noise occurs when cross-multiplying visibilities, which is shown by the two copies of $\textbf{x}$ in Equation \eqref{eq:qhat}. If both copies of $\textbf{x}$ come from the same baseline and time, it can result in power spectrum measurements that are higher than those predicted by the thermal noise of the instrument. One way to avoid this type of noise bias is to avoid cross-multiplying data from the same baselines or days. This ensures that the two quantities that go into a measurement have separate noises that don't correlate with each other.

@@

\begin{equation}
\label{eq:Pin}
P_{in,\alpha} \propto \textbf{e}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{out,\alpha} \equiv \hat{\textbf{p}}_{e,\alpha} &=& \hat{\textbf{p}}_{r,\alpha}-\hat{\textbf{p}}_{x,\alpha} \nonumber \\
&\propto& \textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

@@

One may wonder why $P_{out}$ cannot be computed simply as the weighted power spectrum of $\textbf{e}$ alone, namely $P_{out,\alpha} \propto \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e}$. Expanding Equation \eqref{eq:sigloss}, we see that $P_{out}$ is comprised of multiple terms:

\begin{eqnarray}
\label{eq:crossterm}
P_{out,\alpha} &\propto& (\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&\propto& \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

@@

\begin{eqnarray}
\label{eq:pout_expand}
P_{out, \alpha,\textbf{e} \gg \textbf{x}} &\propto& \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x}.
\end{eqnarray}

We see that our naive expression for $P_{out}$ is the first term in Equation \eqref{eq:pout_expand}, but there are also two additional terms. An initial assumption would be that the cross-terms that involve both $\textbf{e}$ and $\textbf{x}$ should be zero, since foregrounds and the cosmological signal are physically unrelated. In \citet{ali_et_al2015}, this assumption is used, and the first term in Equation \eqref{eq:pout_expand} is directly compared to the input (unweighted) power spectrum to compute signal loss.

However, deeper investigations of these terms reveal that they contain non-negligible power. In fact, foreground modes and signal modes are anti-correlated on average, as derived and explained in \citet{switzer_et_al2015}. The spurious correlation of the two means that we cannot compute signal loss using a signal-only simulation, which would yield greater values for $P_{out}$ and thereby underestimate signal loss. Therefore, in our revised signal loss computation we use the full quantity for $P_{out}$ as defined in Equation \eqref{eq:sigloss}, which subtracts the weighted power spectrum of the data from the weighted power spectrum of data plus EoR.

@@

\begin{eqnarray}
P_{in,\alpha} &\propto& \textbf{e}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e} \\
P_{out,\alpha} &\propto& \textbf{r}^{\dagger}\textbf{C}_{r}^{-1}\textbf{Q}^{\alpha}\textbf{C}_{r}^{-1}\textbf{r} - \textbf{x}^{\dagger}\textbf{C}_{x}^{-1}\textbf{Q}^{\alpha}\textbf{C}_{x}^{-1}\textbf{x} 
\end{eqnarray}

Our signal loss transfer function for one $k$ value is shown in Figure \ref{fig:sigloss_transfercurve} and compares the weighted case ($\textbf{R} = \textbf{C}^{-1}$, left plot) with an unweighted case ($\textbf{R} = \textbf{I}$, right plot). The figures depict a smoothed-out distribution (via kernel density estimators) of our input and output power spectra. To sample the distribution, we run $40$ total injection levels, where the amplitude of \textbf{e} is increased in even increments in log-space and ranges from $P_{in} \sim -10^{13} \, mK^{2} (h^{-1} Mpc)^{3}$ (where one copy of \textbf{e} in Equation \eqref{eq:Pin} is negative and the other is positive) to $P_{in} \sim10^{13} \, mK^{2} (h^{-1} Mpc)^{3}$. For each injection, we perform $20$ bootstraps (over baselines, as explained in Section \ref{sec:Boot}), resulting in a total of $800$ data points on our $P_{in}$ vs. $P_{out}$ grid. It is noted that the transfer function is symmetric for negative and positive $P_{in}$'s, and so we fold all the values into positive ones to increase our signal-to-noise.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve.png}
	\caption{Signal loss transfer functions showing the relationship of $P_{in}$ and $P_{out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Smoothed out distributions (via kernel density estimators) are shown as colored heat-maps for the cases of inverse covariance weighted PAPER-64 data (left) and unweighted data (right). The solid black diagonal line marks a perfect unity mapping, and the solid grey horizontal line denotes the peak of $\hat{\textbf{p}}$, the data distribution. From these plots, it is clear that inverse covariance weighting results in $\sim3$ orders of magnitude of signal loss whereas the unweighted case does not exhibit loss.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

One peculiar aspect of Figure \ref{fig:sigloss_transfercurve} is the fact that at low $P_{in}$ values it appears that we can have signal gain ($P_{out} > P_{in}$). This is unphysical in nature but caused due to the cross-terms dominating in Equation \eqref{eq:pout_expand} once $\textbf{e}$ becomes small. As $\textbf{e}$ increases, we move into a regime where $P_{out} \sim P_{in}$, and then eventually into a regime where $P_{out} < P_{in}$ when $\textbf{e}$ is large enough to destroyed if weighting the data using itself. Although we only show figures for one $k$ value, we note that the shape of the transfer curve is nearly identical for all $k$'s (though we treat each $k$ separately).

@@

The first component is merely the fact that the values of the cross-terms in Equation \eqref{eq:pout_expand} change for different EoR signals, $\textbf{e}$. Since we draw different random $\textbf{e}$'s for every bootstrap, there is a (small) range of $P_{out}$ values that could result for the same $P_{in}$. This randomness is what causes the spread seen in the unweighted case (right plot of Figure \ref{fig:sigloss_transfercurve}), as well as most of the spread in the weighted case (left plot).

@@

Equation \eqref{eq:sense} represents the most sensitive estimate for $\hat{\textbf{p}}(k)$ given our observing parameters. An additional factor of $\sqrt{2}$ is gained in sensitivity when folding our power spectra into $\Delta^{2}(k)$, due to averaging together positive and negative $k$'s.

@@

We fringe-rate filter the noise in the same way as PAPER-64 data, using an optimal fringe-rate filter. Power spectrum results for the noise simulation are shown in Figure \ref{fig:ps_noise}, where the black and grey points represent positive and negative power spectrum values, respectively (with $2\sigma$ error bars and weighting matrix $\hat{\textbf{C}}_{eff}$), the dashed blue line represents the unweighted power spectrum, and the solid green line denotes our $2\sigma$ theoretical noise prediction as calculated by Equation \eqref{eq:sense}. All three are in agreement, validating our analytical thermal noise calculation.

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_noise_reg.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_noise_reg.png}
	\caption{Power spectrum estimate for a noise simulation that mimics the noise level of PAPER-64 data. The weighted power spectrum points and $2\sigma$ errors are show in black and grey (positive and negative values), where we use $\hat{\textbf{C}}_{eff}$ to minimize signal loss. The dashed blue line is the unweighted power spectrum (also $2\sigma$ upper limit). The solid green line is the theoretical noise level prediction as calculated by Equation \eqref{eq:sense}. All three estimates agree.}
	\label{fig:ps_noise}
\end{figure*}

@@

Tailoring window functions is one way to suppress foreground biases. As alluded to in Section \ref{sec:SiglossOverview}, we have a choice for the normalization matrix $\textbf{M}$ in Equation \eqref{eq:phat}. For the analysis of PAPER-64 data, we compute $\textbf{M}$ using the Fisher matrix $\textbf{F}$, defined as:

\begin{equation}
\textbf{F}_{\alpha\beta} = \frac{1}{2} \text{tr} [\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{Q}^{\beta} ]
\end{equation}

@@

In addition to mitigating foreground bias at high $k$'s, two other sources of bias that we actively suppress in the PAPER-64 analysis are noise bias associated with the squaring of thermal noise and noise bias from crosstalk. In order to avoid the former, we filter out certain cross-multiplications when forming $\hat{q}$ in Equation \eqref{eq:qhat}. Namely, the PAPER-64 dataset is divided into two halves: even julian dates and odd julian dates. Our data vectors are then $\textbf{x}_{even, 1}$ for the `even' dataset and baseline group $1$, $\textbf{x}_{odd, 1}$ for the `odd' dataset and baseline group $1$, etc. We only form $\hat{q}$ when the two copies of $\textbf{x}$ come from different groups and baselines, never multiplying `even' with `'even', for example, in order to prevent the squaring of the same thermal noise.

@@

% APPENDIX ---------------------------------------------------------------------------------

@@

The minimum-variance quadratic estimator $\hat{p}_\alpha$ for the $\alpha$th bandpower of the power spectrum is given by \acl{Also need to state in a footnote that for this section only we'll assume without loss of generality that the data are real}
\begin{equation}
\hat{p}_\alpha = \frac{1} {2 \F_{\alpha \alpha} }\x^t \C^{-1} \Q^{\alpha} \C^{-1} \x,
\end{equation}
where
\begin{equation}
F_{\alpha \alpha} \equiv \frac{1}{2} \textrm{tr} \left( \C^{-1} \Q^\alpha \C^{-1} \Q^\alpha \right)
\end{equation}
is the $\alpha$th diagonal element of the Fisher matrix. In our case, however, we do not have \emph{a priori} knowledge of the covariance matrix. Thus, we replace $\C$ with $\Chat$, its data-derived approximation. Our estimator then becomes
\begin{equation}
\label{eq:phatloss}
\hat{p}_\alpha^\textrm{loss} = \frac{1} {2 \F_{\alpha \alpha} }\x^t \Chat^{-1} \Q^{\alpha} \Chat^{-1} \x,
\end{equation}
where
\begin{equation}
\hat{F}_{\alpha \alpha} \equiv \frac{1}{2} \textrm{tr} \left( \Chat^{-1} \Q^\alpha \Chat^{-1} \Q^\alpha \right),
\end{equation}
with the label ``loss" to foreshadow the fact that this will be an estimator with signal loss (i.e., a multiplicative bias of less than unity). We will now provide an explicit demonstration of this by modeling the estimated covariance as
\begin{equation}
\label{eq:ChatDef}
\Chat = (1-\eta) \C + \eta \x \x^t,
\end{equation}
where $\eta$ is parameter quantifying our success at estimating the true covariance matrix. If $\eta = 0$, our covariance estimate has perfectly modeled the true covariance and $\Chat = \C$. On the other hand, if $\eta =1$, then our covariance estimate is based purely on the one realization of the covariance that is our actual data, and we would expect a high level of overfitting and signal loss.

Our strategy for computing the signal loss will be to insert Equation \eqref{eq:ChatDef} into Equation \eqref{eq:phatloss} and to express the resulting estimator $\hat{p}_\alpha^\textrm{loss}$ in terms of $\hat{p}_\alpha$. We begin by expressing $\Chat^{-1}$ in terms of $\C^{-1}$ using the Woodbury identity so that

@@

where $\beta \equiv \eta g/[1+ \eta (g-1)]$ and $\varepsilon \equiv \beta / g$. To proceed, we take the ensemble average of $\hat{p}_\alpha^\textrm{loss}$ to understand its average statistical properties. Inspecting Equation \eqref{eq:phatlossexpanded}, we see that its expectation value is a linear sum of terms of the form $\langle \beta^a h_\alpha^b h_{\alpha \alpha}^c \rangle$, where $a$, $b$, and $c$ are integers. Expanding $\beta$ in powers of $g$, we have

@@

and re-inserting this into Equation \eqref{eq:betahh} yields

@@

With this result, taking the ensemble average of Equation \eqref{eq:phatlossexpanded} is relatively straightforward. One simply replaces every copy of $\beta$ with $\beta_N \equiv \eta N / [1 + \eta (N-1)]$, every copy of $\varepsilon$ with $\varepsilon_N \equiv  \eta  / [1 + \eta (N-1)]$, every copy of $h_\alpha$ with $ \textrm{tr} (\C^{-1} \Q^\alpha )$, and every copy of $h_{\alpha \alpha}$ with $2F_{\alpha \alpha}$. Performing these substitutions and recalling \acl{make sure the reader will indeed be recalling at this point} that $\langle \hat{p}_\alpha \rangle = \langle h_\alpha \rangle / 2 F_{\alpha \alpha}$, we obtain

@@

In general, one's estimate of the covariance will not be expressible in the form given by Equation \eqref{eq:ChatDef}. However, the qualitative point here is that the signal loss can be expressed as a multiplicative factor. Even if $\Chat$ is more complicated than the form used for this toy model, the same mathematical technique of expressing the estimator as a series of higher-point functions, reducing these functions (assuming Gaussianity) to two-point functions, and re-summing any power series will show that the signal loss is a multiplicative factor. \acl{I can express this last bit more eloquently, but it's late and I'm tired and need to eat dinner.}
